---
title: "Social network analysis with statnet"
author: "Kushan De Silva"
date: "April 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Analysis of a friendship network of 1,461 adolescents; a static, undirected, one-mode friendship network of 1,461 vertices.

Three core packages within the statnet suite|network, sna, and ergm are used

exponential-family random graph (ERG) models


```{r}
library("statnet")
data("faux.magnolia.high")
fmh <- faux.magnolia.high
fmh
summary(fmh)

```

network has 1,461 vertices and 974 edges.

```{r}
plot(fmh, displayisolates = FALSE, vertex.cex = 0.7)
table(component.dist(fmh)$csize)

```


There are 524 isolates (which were not included in the visualization), one large component of 439 vertices, and many components in between



```{r}
#network with vertices colored by their grade
plot(fmh, displayisolates = FALSE, vertex.col = "Grade", vertex.cex = 0.7)
```

There is a strong tendency for friendships to form within grade, but the grades themselves do not appear to be cohesive units; rather each grade seems to consist of a number of subgroups which link together

```{r}
fmh.degreedist <- table(degree(fmh, cmode = "indegree"))
fmh.degreedist

summary(fmh ~ degree(0:8))
summary(fmh ~ degree(0:8, "Sex"))
summary(fmh ~ triangle)
summary(fmh ~ edges + triangle)
```


mixing matrix by grade (i.e., the count of relationships cross-tabulated by the grades of
the two actors involved)

```{r}
mixingmatrix(fmh, "Grade")
gr <- fmh %v% "Grade"
table(gr)
```

Fitting an ERG model a.k.a. p* (p-star) model

```{r}
#Bernoulli model
model1 <- ergm(fmh ~ edges)
summary(model1)


```

Density of the network, that is, the fraction of possible edges that are realized, is 0.000913

```{r}
names(model1)
model1$coef
model1$mle.lik

```


Tendency for assortative mixing (a greater probability of individuals forming edges with others of the same race, sex, or grade as themselves; exhibits dyadic independence

Test for a tendency towards assortative mixing that is uniform within each attribute class i.e. there is the same tendency for within-race edges (for example), regardless of which
race one is talking about

```{r}
model2 <- ergm(fmh ~ edges + nodematch("Grade") + nodematch("Race") + + nodematch("Sex"))
summary(model2)

model2$mle.lik
```

```{r}
sim2 <- simulate(model2, burnin = 1e+6, verbose = TRUE, seed = 9)
mixingmatrix(sim2, "Race")
mixingmatrix(fmh, "Race")

plot(summary(fmh ~ degree(0:10)), type = "l", lty = 1, lwd = 2, xlab = "Degree", ylab = "Count")
lines(summary(sim2 ~ degree(0:10)), lty = 2, lwd = 3)
legend("topright", legend = c("Observed", "Simulated"), lwd = 3, lty = 1:2)
c(fmh = summary(fmh ~ triangle), sim2 = summary(sim2 ~ triangle))


```


Identifying model degeneracy


```{r}
model.take1 <- ergm(fmh ~ edges + nodematch("Grade") + nodematch("Race") + nodematch("Sex") + gwesp(0, fixed = TRUE), MCMCsamplesize = 1e+5, maxit = 15, verbose = TRUE, control = control.ergm(steplength = 0.25, seed = 123))

model.take1$coef

model.take2 <- ergm(fmh ~ edges + nodematch("Grade") + nodematch("Race") + nodematch("Sex") + gwesp(0.1, fixed = TRUE), MCMCsamplesize = 1e+5, maxit = 15, verbose = TRUE, control = control.ergm(steplength = 0.25, seed = 123))

model.take3 <- ergm(fmh ~ edges + nodematch("Grade") + nodematch("Race") + nodematch("Sex") + gwesp(0.2, fixed = TRUE), MCMCsamplesize = 1e+5, maxit = 15, verbose = TRUE, control = control.ergm(steplength = 0.25, seed = 123))

c(model.take1$mle.lik, model.take2$mle.lik, model.take3$mle.lik)

model4 <- model.take3
model4$coef


```

```{r}
gof4.deg <- gof(model4 ~ degree, verbose = TRUE, burnin = 1e+5, interval = 1e+5, control.gof.formula (seed = 246))

plot(gof4.deg)
gof4.deg


```


